---
title: "Data Tidying"
author: "Data Science for Public Service"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
---
  
```{r include = FALSE}

setwd("~/DS4PS/New Textbook Sections")

knitr::opts_chunk$set(echo = TRUE)

library(scales)
library(plotly)
library(pander)
library(tidyverse)

```

<br>

# Putting "Tidy" in Tidyverse

The concept of **tidy data** is central to the popular ecosystem of packages
known as the "Tidyverse", hence the name.  

For example, data visualization package `ggplot2` is a key Tidyverse package 
that generally functions correctly only when using tidy data.

Another Tidyverse package, `tidyr`, is designed to turn messy, unruly, or 
generally untidy data into tidy data (and vice versa if one wished!). We'll
learn a bit more about it below.

While familiarization with pivoting data is helpful, it is not the same.
Tidying data involves dismantling to create new rows, columns, and tables. This 
ensures that *all tidy data have a universal format*. 

<br>

#### Thinking Ahead

Before continuing, think a bit on the following quotes in relation to tidy data.

> Happy families are all alike; every unhappy family is unhappy in its own way. (Leo Tolstoy)

> Tidy datasets are all alike but every messy dataset is messy in its own way. (Hadley Wickham)

<br>
<br>

# Defining Tidy Data

When you tidy a room in your home, that's very different from cleaning it. 
When you clean, for example, you might sweep, dust, vacuum, or mop. 
When you tidy, you simply put everything *where it's supposed to be*.

Hence, **"tidy data"** are tabular data consisting of rows (**observations**), 
columns (**variables**), and tables (**types**). More importantly, those 
components must be structured in accordance with three simple tidying rules 
which, sure enough, explain *where everything is supposed to be*: 

1. Each variable forms a column
2. Each observation forms a row
3. Each type of observational unit forms a table

<br>

You can learn more in the original publication "Tidy Data" by Hadley Wickham in the 
**[Journal of Statistical Software](https://www.jstatsoft.org/article/view/v059i10)**.

But what do these rules mean, exactly? Let's find out below.

<br>
<br>

## 1. Each Variable, a Column

Every variable in a tidy dataset must have a corresponding column. For example, 
values from different variables should not exist within the same column, nor
exist as column headers.

<br>

These **untidy data** combine values from different variables in the same column:

```{r echo = FALSE}

data.frame(name_dob = c("Dimitri 1988-04-11", 
                        "Kendra 1999-09-08", 
                        "Monica 1981-10-11"))

```

<br>

These **tidy data** have a column for each variable:

```{r echo = FALSE}

data.frame(name = c("Dimitri",
                    "Kendra",
                    "Monica"),
           dob = c("1988-04-11",
                   "1999-09-08",
                   "1981-10-11"))

```

<br>
<br>

## 2. Each Observation, a Row

Each observation in a tidy dataset must have a corresponding row. One most often
encounters this when dates, which are themselves values, are instead used as 
column headers.

<br>

These **untidy** use years, which are values of variable `Year`, as column headers: 

```{r echo = FALSE}

tibble(Color = c("Red", "Green", "Blue"),
       "2000" = c(2, 4, 6),
       "2001" = c(3, 5, 8),
       "2002" = c(5, 6, 9))

```

<br>

These **tidy data** use two new variables to ensure that all observations now
have their own rows:

```{r echo = FALSE}

tibble(Color = c("Red", "Green", "Blue"),
       "2000" = c(2, 4, 6),
       "2001" = c(3, 5, 8),
       "2002" = c(5, 6, 9)) %>% 
  pivot_longer(cols = "2000":"2002") %>% 
  rename(Year = name,
         Value = value)

```

<br>
<br>

## 3. Each Type, a Table

In a tidy dataset, each *kind* of data must have its own table (Wickham uses the
term "type", though we choose to use "kind"). A table with client salary 
information should not be combined with their treatment records, for instance. 

In such cases, a unique ID is created to separate different *kinds* of data into
distinct tables. Using this key, the tables can be merged as needed.

<br>
<br>

# Tidying Data with 'tidyr'

Package **tidyr** is a Tidyverse package used to help you quickly gather (tidy) 
and spread (untidy) data using functions `pivot_longer()` and `pivot_wider()`, 
respectively. Although, again, this is not pivoting, the naming convention 
refers "long" (tidy) and "wide" (untidy) data. 

Note how literature on tidying data may use many of these terms interchangeably. 

<br>
<br>

## Tidying Demonstration

Recall the basic example of untidy data we used above.

```{r echo = TRUE}

untidy_data <- dplyr::tibble(Color = c("Red", 
                                       "Green", 
                                       "Blue"),
                             "2000" = c(2, 4, 6),
                             "2001" = c(3, 5, 8),
                             "2002" = c(5, 6, 9))

untidy_data

```

<br>

We can use a simple `tidyr` function to make our data "longer" (tidy). It is 
"longer" because we are creating new rows, thus the table becomes taller. 

Here, we use `pivot_long()` and pass the following arguments:

* `data =` accepts our untidy data 
* `cols =` accepts the names of the columns that belong in a new variable
* `names_to =` accepts the name(s) of the new variable(s)
* `values_to =` accepts the names of the value variable(s) created during tidying

<br>

```{r}

tidyr::pivot_longer(data = untidy_data, 
                    cols = "2000":"2002", 
                    names_to = "Year", 
                    values_to = "Value")

```

<br>
<br>

## Basic Untidying Example

Here, we use a subset of the `gapminder` dataset from the **[package of the same name](https://cran.r-project.org/web/packages/gapminder/index.html)**.

This subset includes only the `country`, `continent`, `year`, and `lifeExp` 
(life expectancy). Here's a preview - note that the data are quite tidy!

<br>

```{r echo = FALSE}

library(gapminder)

my_minder <- gapminder %>% 
  select(country, continent, year, lifeExp)

head(my_minder)

```

<br>

Now, we'll **untidy** them by spreading the data using `pivot_wider()`. In this case:

* `names_from =` accepts the variable name containing values used to name new columns
* `values_from =` accepts the variable name containing the values used to populate those columns

<br>

```{r}

my_minder %>% 
  pivot_wider(names_from = year, 
              values_from = lifeExp) %>% 
  head()

```

<br>
<br>

# Further Resources

The world of data tidying is actually quite deep. We encourage you to learn more,
especially about package `tidyr`.

For more information about package `tidyr`, visit the **[official Tidyverse page](https://tidyr.tidyverse.org/)**.

