---
title: "Lab 02 - Identifying Groups"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    theme: cerulean
    highlight: haddock
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 6)

```

<br>

# Data

The following report identifies specific tax parcels in Downtown Syracuse, New York (USA).

View the dataset documentation here: [Syracuse, NY Tax Parcel Data](https://ds4ps.github.io/Data-Science-Class/DATA/syr_parcels.html).

<br>

## Required Packages

The following packages are required for this report and loaded with `library()`.

<br>

```{r}

library(sp)
library(rgdal)
library(geojsonio)                            # Load required libraries

```

<br>

## Import Spatial Polygons

We read in spatial polygons (`"sp"`) with function `geojson_read()` to map 
tax parcel shapes in Syracuse's Downtown area.

<br>

```{r cache = TRUE}

url <- paste0("https://raw.githubusercontent.com/DS4PS/Data",
              "-Science-Class/master/DATA/downtown-syr.geojson")

downtown <- geojson_read(url, what = "sp")    # Load parcel data

plot(downtown,  
     border = "gray60",
     col = "gray80")                          # Visualize parcels

```

<br>
<br>

# Example Solutions

The following examples demonstrate how to use conditions to identify specific parcels.

<br>

### Example: Tax Parcel Acres

Parcels with more than one acre are determined with variable `acres` and condition `> 1`.

* Each value in variable `acres` is tested as greater than one, or `> 1`
* Values in `acres` greater than one become `TRUE`, otherwise `FALSE`
* All `TRUE` and `FALSE` values are stored in object `these`

<br>

```{r}

result <- downtown$acres > 1                  # Each value in 'acres' is tested

```

<br>

All `TRUE` and `FALSE` values, for each parcel, are converted into colors. 

<br>

```{r}

group.colors <- ifelse(test = result,         # 'test =' accepts TRUE or FALSE values
                       yes = "firebrick",     # 'yes =' converts values that are TRUE
                       no = "gray80")         # 'no =' converts values that are FALSE

```

<br>

All tax parcels in `downtown` are mapped with `plot()`.  Each parcel has one of 
two colors per the above code.

<br>

```{r}

plot(downtown,                                # Plots object 'downtown'
     border = NA,
     col = group.colors)                      # Red when TRUE; gray when FALSE

```

<br>
<br>

### Example: Parcels with Single Families

Use variable `landuse` to determine how a tax parcel is used.  Determine all 
possible values with `unique()`. 

<br>

```{r}

unique(downtown$landuse)                      # All unique values in 'landuse'

```

<br>

One of the values in `landuse` is "Single Family".  Therefore, the test is 
whether or not `landuse` is exactly equal to "Single Family".

<br>

```{r}

result <- downtown$landuse == "Single Family" # Test if 'landuse' is "Single Family"

group.colors <- ifelse(test = result,         # Provide TRUE or FALSE test results
                       yes = "firebrick",     # If TRUE, make "firebrick"
                       no = "gray80")         # If FALSE, make "gray80"

plot(downtown,
     border = NA, 
     col = group.colors)                      # Plot with respective colors

```

<br>

# Questions & Solutions

The following questions ask you to map your results, write your answer, or both.
The first question has been completed for you.

<br>
<br>

## Question 1: Vacant Lots

**Question:** *Where are the majority of vacant lots located in the* 
*downtown? Map your results.*

**Note:** *This solution has been provided for you.*

<br>

```{r}

unique(downtown$landuse)                      # Print unique values in 'landuse'

```

<br>

Now, we create our test statement, `== "Vacant Land"`, and store the results in `result`.

<br>

```{r}

result <- downtown$landuse == "Vacant Land"

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

## Question 2: Parking Lots

**Question:** *How many parking lots are in downtown Syracuse? Map your results.*

**Answer:** There are **[x]** parking lots in downtown Syracuse.

<br>

```{r}

# Use function: 'sum()'
# Use variable: 'landuse'

# Assign test (logical) output to object 'result' by replacing 'FALSE'

result <- FALSE

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

## Question 3: New Construction

**Question:** *Where is new construction located in the city?. Map your results.*

**Note:** You may use `>=` (greater than or equal) or `>` (greater than) 1980.

<br>

```{r}

# Use variable: 'yearbuilt'
# Find the number of buildings constructed either after or during and after than 1980

# Assign test (logical) output to object 'result' by replacing 'FALSE'

result <- FALSE

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

**Question:** *What proportion of commercial properties are built since 1980?*

**Answer:** **[X]%** of commercial properties were built since 1980.

<br>

```{r}

# Use variable: 'landuse'
# Use variable: 'yearbuilt'
# Count only commercial properties with buildings constructed during or later than 1980

```

<br>
<br>

## Question 4: Parcels Valued Over $10M

**Question:** *How many parcels have assessed values over $10 million? Map your results.*

**Answer:** **[X]** parcels have assessed values over $10 M.

<br>

```{r}

# Use function: 'sum()'
# Use variable: 'assessedval'

# Assign test (logical) output to object 'result' by replacing 'FALSE'

result <- FALSE

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

## Question 5: Total Tax Delinquent Parcels

**Question:** *How many properties are delinquent on taxes? Map your results.*

**Answer:** **[X]** properties are delinquent on taxes.

<br>

```{r}

# Use function: 'sum()'
# Use variable: 'amtdelinqt'

# Assign test (logical) output to object 'result' by replacing 'FALSE'

result <- FALSE

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

## Question 6: Tax Delinquent Parcels

**Question:** *What proportion of parcels have delinquent tax payments owed?*

**Answer:** **[X%]** of parcels have delinquent tax payments owed.

<br>

```{r}

# Use function: 'mean()'
# Use variable: 'amtdelinqt'

```

<br>
<br>

## Question 7: Tax Delinquent Commercial Properties

**Question I:** *What proportion of commercial properties are delinquent on taxes?*

**Question II:** *What proportion of delinquent tax bills are owed by commercial parcels?*

<br>

**Answer I:** **[X]%** of commercial properties are delinquent on taxes. 

**Answer II:** **[X]%** of delinquent tax dollars are owed by commercial parcels.

<br>

```{r}

# Use function: 'mean()'
# Use variable: 'amtdelinqt'
# Use variable: 'landuse'

# The first answer is tax-delinquent commercial properties over all commercial properties
# The second answer is the tax dollars owed by commercial properties (a subset) over all tax dollars owed

```

<br>
<br>

## Question 8: Tax Delinquent Parcels by Land Use

**Question:** *How many of each land use type are delinquent on taxes? Print a table of your results.*

<br>

```{r}

# Use function: 'table()'
# Use variable: 'amtdelinqt'
# Use variable: 'landuse'

```

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW IF NOT ATTEPTING CHALLENGES** </h4>------------</center>

<br>
<br>

## Challenge I: Parcels Values Over $1M

**Instructions:** *Map all of the land valued at over $1 million per acre.*

<br>

```{r}

# Replace 'FALSE' with your conditional statement

result <- FALSE

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

## Challenge II: Parcels Valued Less Than $500K

**Instructions:** *Map all of the land valued at less than $500,000 per acre.*

<br>

```{r}

# Replace 'FALSE' with your conditional statement

result <- FALSE

group.colors <- ifelse(test = result, 
                       yes = "firebrick", 
                       no = "gray80")

plot(downtown,
     border = NA, 
     col = group.colors)

```

<br>
<br>

## Challenge III: Total Commercial Value

**Instructions:** *What is the total value of all of the commercial parcels in Downtown?*

**Answer:** The total value of all commercial parcels is **[$X]**.

<br>

```{r}

# Your code here

```

<br>
<br>

## Challenge IV: Total Non-Commercial Value

**Question:** *What is the total value of all of the non-commercial parcels in Downtown?*

**Answer:** The total value of all non-commercial parcels is **[$X]**.

<br>

```{r}

# Your code here

```

<br>
<br>

<center>------------<h4> **DELETE THIS LINE & ALL LINES BELOW BEFORE SUBMITTING** </h4>------------</center>

<br>
<br>

# Tips & Tricks

Consider the following tips for completing this assignment.

<br>

### Creating Logical Values

Recall that logical values are `TRUE` and `FALSE`. To produce logical values, we test vaues against a condition, e.g. "greater than".

<br>

```{r}

x <- c(10, 20, 30, 40, 50)    # Creating a vector of values

x > 25                        # Testing each value as greater than 25

```

<br>

### Logical Values as Numeric Values

Recall that logical values are `TRUE` and `FALSE`, representing `1` and `0`, respectively.

<br>

```{r}

as.numeric(TRUE)              # Changing class to "numeric"
as.numeric(FALSE)             # Changing class to "numeric"

```

<br>

### Logical Values & Arithmetic

Because `TRUE` and `FALSE` are actually numeric, we can use `sum()` for total `TRUE` values, e.g.

<br>

```{r}

y <- c(TRUE, TRUE, 
       FALSE, TRUE, 
       FALSE, TRUE)           # Creating a vector of TRUE and FALSE

sum(y)                        # Finding total TRUE values

mean(y)                       # Finding proportion of TRUE values

```

<br>

### Logicals for Subsetting

A subset is a smaller collection of observations (rows) from a larger dataset. 
Create a subset by placing a logical vector in lieu of row positions.

```{r}

index <- downtown$landuse == "Industrial"   # Test if 'landuse' equals "Industrial"

z <- downtown[index, ]                      # Extract only "Industrial" parcels

```

<br>
<br>

# How to Submit

Use the following instructions to submit your assignment, which may vary depending on your course's platform.

<br>

### Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

### Special Instructions

Perform the following depending on your course's platform:

* **Canvas:** Upload both your `.RMD` and `.HTML` files to the appropriate link
* **Blackboard or iCollege:** Compress your `.RMD` and `.HTML` files in a `.ZIP` file and upload to the appropriate link

`.HTML` files are preferred but not allowed by all platforms.

<br>

### Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **Lab-##-LastName.rmd** and **Lab-##-LastName.html**
2. Show both the solution for your code and write out your answers in the body text 
3. Do not show excessive output; truncate your output, e.g. with function `head()`
4. Follow appropriate styling conventions, e.g. spaces after commas, etc.
5. Above all, ensure that your conventions are consistent

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

### Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  

<br>
<br>
<br>