---
title: "Package flexdashboard"
output:
  html_document:
    theme: readable
    highlight: tango
    toc: true
    toc_depth: 2
    css: "../html/lecture-notes.css"
---

```{r include = FALSE}

# setwd("~/DS4PS/New Textbook Sections")

knitr::opts_chunk$set(echo = FALSE)

library(scales)
library(plotly)
library(pander)
library(tidyverse)

```

<br>

# What Is flexdashboard?

Package `flexdashboard` allows you to create Shiny applications using R 
Markdown. 

UI elements and user inputs, as well as all server processing tasks, occur in 
regular R Markdown code chunks, and your script is saved as a `.RMD` file. 

While `flexdashboard` builds have some limitations and significant structural 
differences from Shiny builds, they are very accessible to R Markdown users and 
Shiny learners in general.

<br>
<br>

# The flexdashboard Template

Open a premade `flexdashboard` template in RStudio by clicking on the following:

1. File
2. New File
3. R Markdown...

<br>

In the "New R Markdown" window, use the left navigation menu to select:

1. From Template
2. Flex Dashboard

<br>

This will open a pre-built `.Rmd` `flexdashboard` template with several 
**charts** and **columns** set up. We'll learn more about how important 
**charts** are soon.

<br>
<br>

# Dashboard Components

<br>

## YAML Header

In order to make a `.Rmd` R Markdown script interpretable by `flexdashboard`, 
use the following basic header:

<br>

```{r eval = FALSE, echo = TRUE}

---
title: "My Title"
output:
  flexdashboard::flex_dashboard
---

```

<br>

This is merely the foundation on which many options for customization may be 
added and specified.

<br>
<br>

## Charts

In package `flexdashboard`, **charts** are the meat and potatoes of all
applications. In fact, as soon as you peruse the `flexdashboard` 
**[documentation](https://pkgs.rstudio.com/flexdashboard/articles/flexdashboard.html)**,
you'll quickly see that charts comprise a significant proportion of its contents. 

Charts are initialized and named in your `.Rmd` script with `###` and are 
typically followed by text and an R Markdown code chunk containing your app's 
inputs and outputs.

<br>

```{r eval = FALSE, echo = TRUE}

### Chart A

Some text...

Some code chunk...

### Chart B

Some more text...

Some other code chunk...

### Chart C

Even more text...

Yet another code chunk...

```

<br>
<br>

## Column Orientation

In package `flexdashboard`, dashboards are comprised of **columns**. Within
each column, **charts** appear in the order in which they are coded. 

<br>

We can ensure that `flexdashboard` understands our column layout wishes by 
specificying as much in the YAML header:

<br>

```{r eval = FALSE, echo = TRUE}

---
title: "My Title"
output:
  flexdashboard::flex_dashboard
    orientation: columns
---

```

<br>

We can create a column using the word `Column` and 14 or more dashes (`-`).

<br>

```{r eval = FALSE, echo = TRUE}

Column
--------------
  
### Chart A
  
Code...
  
### Chart B

Code...

```

<br>

Column widths may be modified with `Column {data-width = <number>}` and the
width of all columns must equal 1,000!

<br>

```{r eval = FALSE, echo = TRUE}

Column {data-width = 250}
--------------
  
### Chart A
  
Code...

Column {data-width = 250}
--------------
  
### Chart B
  
Code...

Column {data-width = 500}
--------------
  
### Chart C
  
Code...

```

<br>
<br>

## Row Orientation 

Rather than **column** orientation, note that we can specify **row** orientation
using the YAML header.

<br>

```{r eval = FALSE, echo = TRUE}

---
title: "My Title"
output:
  flexdashboard::flex_dashboard
    orientation: rows
---

```

<br>
<br>

# Dashboard Tabsets

<br>

## Chart Tabsets

In `flexdashboard`, **tabsets** allow your **charts** to have multiple dropdown
tabs which may contain additional charts or text. 

Simply add `Column {.tabset}` to a column header and all ensuing charts will 
become named dropdown menu items for the existing chart space.

<br>

```{r echo = TRUE, eval = FALSE}

Column {.tabset}
--------------
  
### Dropdown Chart 1
  
Some code...
  
### Dropdown Chart 2

Some code...
  
### Dropdown Chart 3

Some code...

```

<br>
<br>

## Page Tabsets 

**Tabsets** also allow your **pages** to have multiple tabs which may contain 
additional columns, charts, and tabsets. 

In order to create a page: 

1. On the first line, write the name of your page
2. On the secodn line, place 16 or more equal signs (`=`)

<br>

Simply add `Page {data-navmenu = Details}` to a page header and all ensuing 
charts will become named dropdown menu items for the existing dashboard page 
space, accessible in the top navigation bar.

<br>

```{r echo = TRUE, eval = FALSE, lang = "HTML"}

Page {data-navmenu = <"Sub-Page Name">}
================
  
### Dropdown Chart 1
  
Some code...
  
### Dropdown Chart 2

Some code...
  
### Dropdown Chart 3

Some code...

```

<br>
<br>

# Shiny Implementation

In order to add interactivity using Shiny, you'll need to modify the **YAML header**
with `runtime: shiny`.

<br>

```{r eval = FALSE, echo = TRUE}

---
title: "My Title"
output:
  flexdashboard::flex_dashboard
    orientation: columns
runtime: shiny
---

```

<br>

Once this is done, we're able to add typical reactive elements one sees in Shiny. 

<br>

## Adding a Sidebar

If you've made your `flexdashboard` report interactive, you'll likely want a
sidebar where users can interact with filters and other options. 

In this case, create a new column, albeit with `.sidebar` applied:

<br>

```{r echo = TRUE, eval = FALSE}

Column {data-width = 250 .sidebar}
--------------

```

<br>
<br>

## Go Forth with Shiny

Using this novel approach to Shiny dashboard development, continue to learn
about Shiny reactivity as, beyond these basic structural differences, 
package `flexdashboard` operates in the same way under the hood. 

<br>
<br>
