---
output:
  html_document:
    theme: readable
    highlight: tango
    css: '../html/lecture-notes.css'
---
  
```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scales)
library(plotly)
library(tidyverse)
```

# Data Frames

**Data frames** are tabular data objects in R, meaning that they are comprised 
of rows and columns. Unlike **matrices**, **data frames** can have variables of 
different classes, such as numeric, character, logical, and datetime. 

Odds are, you've already seen and worked with several in R. Classic practice
data frames include `mtcars` and `iris`. Let's print one the old-fashioned way.

<br>

```{r}

mtcars %>% 
  head() %>% 
  print()

```

<br>
<br>

## Class 'data.frame'

Using function `class()` on a **data frame** object reveals class `data.frame`.

<br>

```{r}

class(mtcars)

```

<br>

You can recast other tabular objects, like **matrices**, into class `data.frame`
using `as.data.frame()` or `dplyr::as_tibble()`.

<br>

```{r}

class(volcano)

volcano_df <- as.data.frame(volcano)

class(volcano_df)

```

<br>

Note that some objects may be of multiple classes, one of which may often be of
class `data.frame`:

<br>

```{r}

mtcars_df <- dplyr::as_tibble(mtcars)

class(mtcars_df)

```

<br>
<br>

## Anatomy

Each **data frame** is composed of one or more **vectors** of equal length. 
Recall that the following is a vector of length 5 (it has 5 elements):

<br>

```{r}

lyrics <- c("Linden", "Arden", "stole", "the", "highlights.")

```

<br>

We can create another vector of numbers for the order of `lyrics`:

<br>

```{r}

order <- 1:5

```

<br>

With function `data.frame()`, we can combine the above vectors to create a
tabular data object of class `data.frame`. 

<br>

```{r}

created_df <- data.frame(order, 
                         lyrics)

class(created_df) 

print(created_df)

```

<br>

You can even create a custom **data frame** entirely from scratch, like 
these top five artists using **[data from The Pudding on lyrical diversity](https://pudding.cool/projects/vocabulary/index.html)**:

<br>

```{r}

most_diverse_lyrics <- data.frame(artists = c("Aesop Rock", 
                                              "Busdriver", 
                                              "Jedi Mind Tricks",
                                              "GZA",
                                              "Wu-Tang Clan"),
                                  words = c(7879,
                                            7324,
                                            6424,
                                            6390,
                                            6196))

print(most_diverse_lyrics)

```

<br>
<br>

## Extraction & Reassembly

We may dismantle a **data frame** by its constituent vectors using the 
**extraction operator** or `$`. This is the most straightforward and common 
approach to extracting a variable from a **data frame**.

```{r}

artists_extr <- most_diverse_lyrics$artists     # Extract 'artists'

words_extr <- most_diverse_lyrics$words         # Extract 'words'

df_extr <- data.frame(artists = artists_extr,
                      words = words_extr)       # Recreate data frame, rename

most_diverse_lyrics == df_extr                  # Are the data frames equal?

```

<br>
<br>
