---
title: "Introductory Programming"
output:
  html_document:
    theme: readable
    highlight: tango
    toc: true
    toc_depth: 2
    css: "../html/lecture-notes.css"
---

```{r include = FALSE}

# setwd("~/DS4PS/New Textbook Sections")

knitr::opts_chunk$set(echo = TRUE)

library(scales)
library(tidyverse)

```

# On Conditional Programming

We've already seen how to use conditional statements with comparators like
`==`, `>`, and `<` in order to create subsets with our data. 

In the same way,
**[conditional programming](https://en.wikipedia.org/wiki/Conditional_(computer_programming))**
uses conditional statements to control how certain scripted instructions are
handled, like whether:

* A specific approach should be used to analyze data
* Some plot should be rendered using a particular function
* Date should be grouped by certain variables

<br>
<br>

## If-Then Statements

In R and many other languages, an 
**[If-Then Statement](https://en.wikipedia.org/wiki/Conditional_(computer_programming)#If-then(-else))**
uses a conditional statement to determine how to proceed with additional 
instructions. Its basic format is as follows:

<br>

```{r eval = FALSE, echo = TRUE}

if (<conditional>){
  
  <instructions>
  
}

```

<br>

For example, suppose we want to use dataset `ggplot2::economics` to determine 
average U.S. unemployment across all dates. Moreover, we would like to classify 
each date as either "Above" or "Below" average.

The following creates our average unemployment value: `avg_unemp`.

<br>

```{r}

library(ggplot2)

avg_unemp <- mean(economics$unemploy, 
                  na.rm = TRUE)

print(avg_unemp)
  
```

<br>

Next, we can create a simple script with our **If-Then Statement**, where `x` 
can represent any date's unemployment value. 

If the date represented by `x` has unemployment (`unemploy`) less than the 
average (`avg_unemp`), our script will return "Below Average". 

<br>

```{r}

x <- economics$unemploy[economics$date == "1989-08-01"]

if (x < avg_unemp){
  
  print ("Below Average")
  
}

```

<br>

Note that we can create an additional **If-Then Statement** to identify whether
`unemploy` is also "Below Average".

<br>

```{r}

x <- economics$unemploy[economics$date == "2012-01-01"]

if (x < avg_unemp){
  
  print ("Below Average")
  
}

if (x > avg_unemp){
  
  print ("Above Average")
  
}

```

<br>
<br>

## Using "Else"

What happens when none of our conditions are met but we still wish to provide 
instructions for a "catch all" category? In this case, we use **Else**, a simple
modification used to create **If-Then-Else Statements**.

The format is straightforward:

<br>

```{r echo = TRUE, eval = FALSE}

if (<condition>){
  
  <instructions if TRUE>
  
} else {
  
  <instructions if FALSE>
  
}

```

<br>

Here, we'll combine our two **If-Then Statements** above that classify 
unemployment values:

<br>

```{r}

x <- economics$unemploy[economics$date == "2015-04-01"]

if (x < avg_unemp){
  
  print ("Below Average")
  
} else {
  
  print ("Above Average")
  
}

```

<br>
<br>

# Iterating with Conditionals

Note that we've only applied our **If-Else Statements** to individual values. 
What if we want to iterate over multiple values, or even an entire variable, to 
evaluate whether each of them pass our conditions?

When we describe **iterating** in this context, we mean *literally processing* 
*some code from one value to the next*. We explore **iteration** in this sense
when discussing **For Loops** below.

<br>
<br>

## Function 'ifelse()'

Using function `ifelse()` is perhaps the easiest way to iterate over any number 
of values and evaluate whether those values pass one or more conditions.

The format is straightforward:

<br>

```{r eval = FALSE, echo = TRUE}

ifelse(test = <conditional>, 
       yes = <value if TRUE>, 
       no = <value if FALSE>)

```

<br>

Do check out the documentation with `?ifelse` or `help(ifelse)`.

Suppose we would like to recreate our average employment classification with 
our above examples:

<br>

```{r}

print(c(x, avg_unemp))

ifelse(test = x < avg_unemp, 
       yes = "Below Average", 
       no = "Above Average")

```

<br>

Importantly, function `ifelse()` not only evaluates conditional statements for a 
single value, it can evaluate all values in an entire variable!

<br>

```{r}

ifelse(test = economics$unemploy < avg_unemp, 
       yes = "Below Average", 
       no = "Above Average") %>% 
  head()

```

<br>

In effect, this allows us to create new variables entirely:

<br>

```{r}

economics$category <- ifelse(test = economics$unemploy < avg_unemp, 
                             yes = "Below Average", 
                             no = "Above Average")

head(economics)

```

<br>

We can achieve similar ends the Tidyverse way using `mutate()` and `ifelse()`:

<br>

```{r}

library(dplyr)

economics <- economics %>% 
  mutate(category = ifelse(test = economics$unemploy < avg_unemp, 
                             yes = "Below Average", 
                             no = "Above Average"))

head(economics)

```

<br>

Do check out function `case_when()` from package `dplyr` for more advanced
classification examples.

<br>
<br>

## For Loops

**[For Loops](https://en.wikipedia.org/wiki/For_loop)** are 
**[Control Flow Statements](https://en.wikipedia.org/wiki/Control_flow)** that 
allow you to iterate, that is, *step from one value to the next in a series*. 

<br>
<br>

### Anatomy of a For Loop

Let's continue with the example of `economics$unemploy`. 

<br>

#### Iterators

Firstly, you need an **iterator**. Luckily, it only requires 
making a name. You may need to name more than one. Usually, you'll see: 

* `i`, `j`, `k` ... 
* `i`, `ii`, `iii` ...
* `a`, `b`, `c` ...

You may even see something more specific like `students`.

**In our case,** we'll name it `iter`.

<br>

#### Sequence

Secondly, you need a **sequence**, which is a series of identifiers used as 
stepping stones for each iteration. The length of the sequence should match 
the length of the number of values over which you wish to iterate.

Let's check out the number of dates in `economics`:

<br>

```{r}

length(economics$unemploy)

```

<br>

**In our case,** our sequence will need to have 
`r comma(length(economics$unemploy))` values, one for each value in variable
`unemploy`.

<br>

```{r}

unemp_seq <- 1:length(economics$unemploy)

head(unemp_seq)

```

<br>
<br>

### For Loop Structure

The default format for a **For Loop** is as follows:

<br>

```{r eval = FALSE}

for (<iterator> in <sequence>){
  
  <intructions>[<iterator>]
  
}

```

<br>

**In our case,** recall that `unemp_seq` is the sequence 
`1:length(economics$unemploy)`. We'll use iterator `iter`.

<br>

```{r eval = FALSE}

for (iter in unemp_seq){
  
  ...
  
}

```

<br>
<br>

### For Loop Logic

This part is a bit tricky, only in that it takes a moment to click. 

Your **iterator**, `iter`, will take on or represent *each number* in your 
sequence for *each iteration*.

If the first number in your sequence is "1", your iterator, `iter`, then 
represents 1. The second number is "2" and `iter` represents 2. The third 
number is "3" and `iter` represents 3. What's that look like, you ask?

<br>

```{r eval = FALSE}

iter <- 1
my_data[iter, ]

iter <- 2
my_data[iter, ]

iter <- 3
my_data[iter, ]

```

<br>

This in effect allows iterating over each row in a data frame or each value 
in a vector. 

<br>

```{r eval = FALSE}

my_date[1, ]
my_date[2, ]
my_date[3, ]

```

<br>
<br>

#### Demonstration

Using out unemployment data, `ggplot2::economics$unemploy`, let's create a 
simple **For Loop** that will iterate over total unemployment for each date and 
categorize whether it will be above or below average, similar to our above 
example.

<br>

```{r}

library(ggplot2)

avg_unemp <- mean(economics$unemploy, 
                  na.rm = TRUE)                 # Calculate average

unemp_seq <- 1:length(economics$unemploy)       # Create sequence

economics$iteration <- NA                       # Initialize new variable

for (iter in unemp_seq){                        # Name iterator and sequence
  
                                                # Instructions for iteration
  
  economics$iteration[iter] <- ifelse(test = economics$unemploy[iter] > avg_unemp, 
                                      yes = "Above", 
                                      no = "Below")
  
}

head(economics)

```

<br>
<br>
