---
title: "Matrices & Lists"
author: "Data Science for Public Service"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
---
  
```{r include = FALSE}

setwd("~/DS4PS/New Textbook Sections")

knitr::opts_chunk$set(echo = TRUE)

library(scales)
library(plotly)
library(tidyverse)

```

# Enter the Matrix

In R, a **matrix** is a tabular data object that is similar to a **data frame**. 

However, unlike the latter, a **matrix** cannot have more than one class of 
data - it must be either entirely of class numeric, logical, character, etc. 

We can create one using function `matrix()`.

```{r}

matrix(1:10, nrow = 5)

```

<br>
<br>

## Class 'matrix'

We can recast **data frames** to objects of class `matrix` using `as.matrix()`.

```{r}

head(iris3)

class(iris3)

as.matrix(x = iris3) %>% 
  class()

```

<br>
<br>

## Will This Be On the Test?

What we won't do often with **matrices**, however, is *use them*. They are 
valuable in certain use cases of which we will likely encounter few. 

<br>
<br>

## Redeeming Qualities

The `matrix` object is still pretty cool. Take a look at the 
`volcano` object from package `plotly`. It's a matrix that creates a square 
grid, with each containing an 
**[elevation level of a New Zealand volcano](https://plotly.com/r/3d-surface-plots/)**. 

<br>

```{r fig.align = "center"}

library(plotly)

volcano[1:6, 1:10] %>% 
  print()

plot_ly(z = ~volcano) %>% 
  add_surface()

```

<br>
<br>

# Lists

Compared to the **matrix**, **lists** are far more common and generally 
extremely useful, though we'll not explore **lists** too deeply in any 
introductory modules.

As such, we will lightly go over **lists** conceptually, as you will encounter 
them, often unwittingly, throughout your journey in R.

<br>
<br>

## What's a List?

In R, a **list** is the next evolutionary phase of the **data frame**.

A **list** is similar to a **data frame** in that they both may contain a 
medley of different classes. In fact, you'll often hear that a **data frame** 
is simply a primitive **list**, which is mostly true.

Here, we can create a list of names vectors using function `list()`:

<br>

```{r}

library(ggplot2)
library(gapminder)

my_list <- list(statement = c("You", "should", "meet", "my", "father."),
                inquiry = c("Your", "father's", "still", "alive?"),
                response = c("No,", "but", "you", "should", "meet", "him."))

```

<br>
<br>

## Accessing Lists, Briefly 

Accessing the contents of **lists** is very challenging at first, but it is 
extremely rewarding. We ecourage all R learners to explore them as early as 
they are comfortable. 

<br>

```{r}

my_list[1]

my_list[[2]]

my_list[[3]][3:6]

```

<br>
<br>

## Advantages of Lists 

**Lists** can hold far more than **vectors** of unequal length. 
**Lists** can hold entirely different **matrices**, multiple **data frames**, 
and even other **lists**! Note that the following list, `my_list`, is created 
with a combination of custom vectors, built-in R data frames, package-built 
data frames, and a numeric matrix!

<br>

```{r}

library(plotly)
library(ggplot2)
library(gapminder)

my_list <- list(some_numbers = 1:10,
                some_letters = letters[1:10],
                economics = economics,
                diamonds = diamonds,
                faithful = faithful,
                volcano = volcano,
                mtcars = mtcars,
                gapminder = gapminder)

my_list$diamonds %>% 
  head() %>% 
  print()

my_list$gapminder %>% 
  head() %>% 
  print()

```

<br>

**Lists** are awesome.

<br>
<br>

## Lists Are Everywhere

One thing to note about **lists** is that, despite our relatively little use of 
them in most modules, many objects you'll use in R will actually be **lists**,
though they are usually called something more specific.

Below, we make a "graphical object", or "grob", using package `ggplot2`. In
functionality, these are and may be accessed as **lists**.

<br>

```{r}

library(ggplot2)

my_plot <- qplot(x = economics$date, 
                 y = economics$pop) +
  theme_minimal()

my_plot

my_plot$labels                    # Access the internals of 'my_plot' with '$'

unlist(my_plot)[1:3]              # Alternatively, use 'unlist()' or like functions

```

<br>
<br>
