---
title: "Your Topic Title"
author: "Your Name"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: yes
    toc_float: yes
---

```{r include = FALSE}

# SET GLOBAL KNITR OPTIONS

knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10, 
                      fig.height = 8)
```

<br>
<br>

# Introduction

This code through explores...

<br>

## Content Overview

Specifically, we'll explain and demonstrate...

<br>

## Why You Should Care

This topic is valuable because...

<br>

## Packages

We will use the following packages for the vignette.

```{r}
library(pander)
library(kableExtra)
```

## Learning Objectives

Specifically, you'll learn how to...

<br>
<br>

# Body Title

Here, we'll show...

<br>

## Further Exposition

This is based/expanded on the theory/work/extension of... 

<br>

## Basic Example

A basic example shows how...

```{r}

# Some code

```

<br>

## Advanced Examples

More specifically, this can be used for...

```{r}

# Some code

```

<br>

What's more, it can also be used for...

```{r}

# Some code

```

<br>

Most notably, it's valuable for...

```{r}

# Some code

```

<br> 
<br>


# Works Cited

This code through references and cites the following sources:

<br>

* Author (Year). Source I. [Hyperlink Text](https://www.google.com)

* Author (Year). Source II. [Hyperlink Text](https://www.google.com)

* Author (Year). Source III. [Hyperlink Text](https://www.google.com)

<br>

## Additional Resources

Learn more about [package, technique, dataset] with the following:

<br>

* Resource I [Hyperlink Text](https://www.google.com)

* Resource II [Hyperlink Text](https://www.google.com)

* Resource III [Hyperlink Text](https://www.google.com)

<br>
<br>


<br>