---
title: "Lab 06: Dataset Joins"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: yes
    toc_float: no
    self-contained: true
    css: '../html/lab-instructions.css'
    include:
      after_body: '../html/footer.html'
---

```{r include=FALSE}

library(readr)
library(dplyr)
library(tidyr)
library(tibble)
library(pander)

knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 10)     # Global knitr options

options(scipen = 999)                     # Disable sci. notation

`%!in%` <- Negate(`%in%`)                 # Custom filter function

url <- "https://www.onetcenter.org/dl_files/database/db_26_3_text/"

dbs <- c("Skills",
         "Knowledge",
         "Abilities",
         "Interests",
         "Job%20Zones",
         "Work%20Values",
         "Work%20Styles",
         "Occupation%20Data",
         "Alternate%20Titles",
         "Technology%20Skills",
         "Job%20Zone%20Reference",
         "Education%2C%20Training%2C%20and%20Experience",
         "Education%2C%20Training%2C%20and%20Experience%20Categories")

txt <- paste0(url, dbs, ".txt")

all <- lapply(txt, read_delim)

names(all) <- gsub(pattern = "Education%2C%20Training%2C%20and|%20",
                   replacement = "",
                   x = dbs)

for(i in 1:length(all)) assign(names(all)[i], all[[i]])   # Create objects

rm(all)

```

# Background

In this hypothetical scenario, pretend that you are a student currently learning
the R language. Take your time and really get into the role. 

We're going to use public data from the U.S. Department of Labor in order to 
lightly research various R-related occupations, the values and interests of 
R-using occupational incumbents, the labor market value of the R language, and 
we conclude by predicting your future salaries based on your work values!

<br>

## Data: O*NET & SOC Codes

O*NET, or the [Occupational Information Network](https://www.onetonline.org/), 
is the admirable pursuit by the Department of Labor to consolidate the 
classification of approximately one thousand kinds of jobs. If that seems 
like a lot, it comes from the even more voluminous DOT 
([Dictionary of Occupational Titles](https://en.wikipedia.org/wiki/Dictionary_of_Occupational_Titles)), 
which defined over 13,000 jobs since 1938. To do this, all federal agencies use 
a unique identifier for each job type, known as a SOC or 
[Standard Occcupational Classification](https://www.bls.gov/soc/). SOC codes,
typically variable `O*NET-SOC Code`, most often become the unique keys necessary 
to joining our tables!

<br>

The [O*NET Resource Center](https://www.onetcenter.org/database.html) contains 
all [individual tables](https://www.onetcenter.org/database.html#individual-files) 
used to power My Next Move and other platforms. Tables are organized according 
to myriad dimensions, seen here:

<br>

```{r echo = FALSE, fig.align = "center", out.width = "75%"}

knitr::include_graphics("https://www.onetonline.org/image/contentmodel.svg")

```

<br>

## Exploring O*NET Tables

Let's look at a few tables to get to know the data.

<br>

### Occupation Titles

Read in one of the highest-level tables, 
[Occupation Data](https://www.onetcenter.org/dictionary/26.3/excel/occupation_data.html),
which contains:

1. Occupation SOC Code
2. Occupation Title
3. Occupation Description

<br>

```{r echo = TRUE}

url <- "https://www.onetcenter.org/dl_files/database/db_26_3_text/Occupation%20Data.txt"

occupations <- read_delim(url, show = FALSE)

occupations %>% 
  head(3) %>%          # Show first 3
  pander()

```

<br>

### Occupation Technologies

Now, let's take a look at a more sophisticated table, 
[Technology Skills](https://www.onetcenter.org/dictionary/26.3/excel/technology_skills.html),
which contains:

1. Occupation SOC Code
2. Technology Title
3. Technology Example
4. UNSPSC Commodity Code
5. UNSPSC Commodity Title
6. "Hot" Technology

<br>

Here, the first 6 observations all have the same SOC Code. That's because
"Technology Skills" lists all technologies associated with the occupation. Our 
preview shows the first 6 technologies important for occupation SOC code
`11-1011.00`, or that of "Chief Executives".

<br>

```{r echo = TRUE}

url <- "https://www.onetcenter.org/dl_files/database/db_26_3_text/Technology%20Skills.txt"

technologies <- read_delim(url, show = FALSE)

technologies %>% 
  head(6) %>%          # Show first 6
  pander()

```

<br>

## Joining Tables

Recall that SOC Codes are unique identifiers for each occupation and are most 
often used fully, or in part, as merge keys. 

<br>

We can merge these using `dplyr` function `left_join()`, specifying 
argument __by =__:

```{r eval = FALSE}

occupations %>% 
  left_join(technologies, 
            by = `O*NET-SOC Code`)

```

<br>

Alternatively, we can use base R function `merge()`:

```{r eval = FALSE}

merge(x = occupations, 
      y = technologies, 
      by = "O*NET-SOC Code")

```

<br>

In fact, if we simply use `left_join()` without identifying a merge column, it 
automatically detects merge keys with common names and notifies us in the 
console.

```{r}

occupations %>% 
  left_join(technologies) %>% 
  select(Title, `Commodity Title`) %>% 
  head(100)

```

<br>

## Insights with Joins

Now that we're capable of joining O*NET database tables, we can begin to glean 
our first insights. For example: 

1. How many technologies are associated with each occupation? 
2. Which occupations require the most diverse array of technologies? 

<br>

We can capture this quickly because our occupations-to-technologies merge is
**one-to-many**, meaning that one record in one table corresponds to multiple 
records in another, adjoining table. Hence, if we count the number of new rows 
created by merging occupation titles and their many technologies, a simple row 
count will sum up to total unique technologies by occupation. 

<br>

```{r}

occupations %>% 
  left_join(technologies) %>% 
  group_by(Title) %>% 
  summarize(Technologies = n()) %>% 
  arrange(-Technologies) %>% 
  head(10)

```

<br>

# Instructions

## Lab Template

Downlaod the lab **template**: 

<a href='../../templates/' target="_blank">
<button type="button" class="button"> TEMPLATES </button>
</a>

*You may have to right-click on the file and “save as” depending upon your browser.*

Remember to name your file: **lab-##-lastname.rmd**

<br>

## Data

The following objects have been created for you from the [O*NET 26.3 Database](https://www.onetcenter.org/database.html). 

1. `Abilities` measures innate or natural abilities of occupational incumbents
2. `AlternateTitles` lists over 53,000 alternate names for O*NET occupations
3. `Experience` measures various kinds of experience, albeit encoded
4. `ExperienceCategories` decodes measured experience
5. `Interests` measures occupations with the Holland Code (RIASEC) Test  
6. `JobZoneReference` decodes classes of career development
7. `JobZones` broad classes of career development for each occupation; encoded
8. `Knowledge` defines areas of expertise and importance
9. `OccupationData` lists the occupation SOC, title, and description
10. `Skills` measures learned abilities of occupational incumbents
11. `TechnologySkills` lists all relevant technology skills
12. `WorkStyles` measures various performance indicators
13. `WorkValues` measures important occupation qualities 

<br>
<br>
<br>



# Part 1: Basic Join Challenges

The following challenges require no more then 2-3 tables and use the same column 
names across all tables. Use basic `dplyr` verbs to answer specific questions.

<br>

## (1) Question 

Which 10 occupation titles (`OccupationData$Title`) have the highest number of "hot technologies" (`TechnologySkills$HotTechnology`)?

```{r}

# Code

```

<br>

## (2) Question: 

Which 10 occupation titles (`OccupationData$Title`), with morethan 15 unique technologies, have the highest proportion (%) of "hot technologies" (`TechnologySkills$HotTechnology`)?

```{r}

# Code

```

<br>

## (3) Question 

Which unique occupation titles (`OccupationData$Title`) have 
"R" listed as a technology example (`TechnologySkills$Example`)? 

```{r}

# Code

```

<br>

## (4) Question 

Which alternate occupation titles (`AlternateTitles`) have "R" 
listed as a technology example (`TechnologySkills$Example`)? 

```{r}

# Code

```

<br>

## (5) Question 

Which top 25 occupation titles (`OccupationData$Title`) have 
the highest level of "Independence" (`Element Name`) according to `WorkValues`?

```{r}

# Code

```

<br>

## (6) Question

Which top 25 occupation titles (`OccupationData$Title`) have 
the highest level of "Persistence" (`Element Name`) according to `WorkStyles`?

```{r}

# Code

```

<br>

## (7) Question 

According to Holland Codes, "Realistic" describes people who 
like to work with things, requiring "motor coordination, skill, and strength".
Which top 20 occupation titles (`OccupationData$Title`) have the highest rating 
in "Realistic" scores (`Element Name`) according to `Interests`?

```{r}

# Code

```

<br>

## (8) Question

Again, per Holland Codes, "Investigative" describes people who 
like to work with ideas, preferring observation over action and facts over 
feelings. Which top 20 occupation titles (`OccupationData$Title`) have the 
highest rating in "Investigative" scores (`Element Name`) according to `Interests`?

```{r}

# Code

```

<br>

## (9) Question

Job Zones loosely classify how advanced one typically must be 
as an occupational incumbent. How many unique occupation titles 
(`OccupationData$Title`) are classified in each `Job Zone` (`JobZones`)? 

```{r}

# Code

```

<br>

## (10) Question

Of occupation titles (`OccupationData$Title`) that list "R" 
as the `Example` in `TechnologySkills`, what proportion is in each Job Zone 
(`JobZones`)?

```{r}

# Code

```

<br>
<br>
<br>

# Part 2: Joins & Salary Predictions

The following challenges require 3-4 merged tables and incorporate data from the 
U.S. Bureau of Labor Statistics' 2021
[Occupational Employment and Wage Statistics](https://www.bls.gov/oes/current/oes_nat.htm)
which, like other federal agencies, uses unique SOC codes for each occupation.

<br>

The following commands will read in BLS data and rename and reformat the columns 
to match O*NET.

```{r, message = FALSE, warning = FALSE}

url <- paste0("https://raw.githubusercontent.com/DS4PS/ays-r-cod",
              "ing-sum-2022/main/labs/bls_occupation_salaries.csv")

Salaries <- read_csv(url, 
                     col_select = c("O*NET-SOC Code" = "OCC_CODE", 
                                    "Title" = "OCC_TITLE", 
                                    "Median" = "A_MEDIAN")) %>% 
  mutate(`O*NET-SOC Code` = gsub(x = `O*NET-SOC Code`, 
                                 pattern = "$", 
                                 replacement = "\\.00"),
         Median = gsub(x = Median, 
                       pattern = ",", 
                       replacement = ""),
         Median = as.numeric(Median))

```

<br>

**Note:** There is one key distinction between O*NET and BLS SOC codes. The 
[2019 O*NET Taxonomy](https://www.onetcenter.org/taxonomy.html) features even 
more detailed profiles than typical federal agencies. Some more detailed O*NET 
occupation profiles will not successfully merge with BLS occupations.

<br>

## (1) Question

Join the O*NET `OccupationData` and BLS `Salaries` tables. 
Then, filter out any occupations with an `NA` value in variable `Median`.
Store this new table as object `Common`. How many occupations have
salary data available from BLS?


```{r}

# Code
  
```

Use basic `dplyr` verbs to answer each question.

<br>

## (2) Question

Join new object `Common` with object `Interests`. Note that 
each interest is in `Element Name` and each interest rating (on a 7-point scale) 
is listed in `Data Value`. In effect, we can now associate each "interest" (e.g. 
artistic, social, enterprising) with median annual salaries.

Use the `lm()` function with formula `Median ~ Element Name` to print the 
coefficients of a simple linear model. Which two element names (not "High 
Points") are most associated with higher salaries?

```{r}

# Code

```

<br>

## (3) Question

Now join `Common` with `TechnologySkills`. Filter occupations 
to only include the following technologies in `Example`, which have been 
provided for you in object `tech`: 

* R
* C
* Python
* JavaScript
* Microsoft Excel
* Spreadsheet software
* SAS statistical software
* Hypertext markup language HTML

Group by technologies (`Example`) and determine the following summaries:

* Average value for `Median` (annual salary)
* Count for total occupations with said `tech` 

```{r}

tech <- c("R",
          "C++",
          "Python",
          "JavaScript",
          "Microsoft Excel",
          "Spreadsheet software",
          "SAS statistical software",
          "Hypertext markup language HTML")

# Code

```

<br>

## (4) Question 

Please rate your own work values by providing a score from 
`1` to `7` for the following value categories.

* Achievement
* Working Conditions
* Recognition
* Relationships
* Support
* Independence 

A "7" indicates highest importance. A "1" indicates lowest importance. 
A default of `3.5` has been provided as an example.

```{r echo = TRUE}

my_values <- tibble::tibble("Achievement"        = 3.5, 
                            "Working Conditions" = 3.5,
                            "Recognition"        = 3.5,
                            "Relationships"      = 3.5,
                            "Support"            = 3.5,
                            "Independence"       = 3.5)

```

<br>

## (5) Question

Join objects `Common` and `WorkValues` and filter the 
following values from variable `Element Name`:

* "First Work Value High-Point"
* "Second Work Value High-Point"
* "Third Work Value High-Point"

Then, use `select()` on the following variables:

* "O*NET-SOC Code"
* "Element Name" 
* "Data Value" 
* "Median"

Pipe (`%>%`) your output into the following `tidyr` function:

```{r eval = FALSE, include = TRUE}

# Code

tidyr::pivot_wider(names_from = "Element Name", 
                   values_from = "Data Value")

```

<br>

Lastly, use function `lm()` and `predict()` to estimate your salary based on 
your work values! 

Make sure to set `eval = TRUE` for the following to run:

```{r eval = FALSE, include = TRUE}

my_model <- lm(formula = Median ~ 
                 Achievement + 
                 `Working Conditions` + 
                 Recognition + 
                 Relationships + 
                 Support + 
                 Independence, 
               data = you_data)

predict(object = my_model, newdata = my_values)

```

<br>





# How to Submit

Use the following instructions to submit your assignment, which may vary depending on your course's platform.

<br>

## Knitting to HTML

When you have completed your assignment, click the "Knit" button to render your `.RMD` file into a `.HTML` report.

<br>

## Special Instructions

Perform the following depending on your course's platform:

* **Canvas:** Upload both your `.RMD` and `.HTML` files to the appropriate link
* **Blackboard or iCollege:** Compress your `.RMD` and `.HTML` files in a `.ZIP` file and upload to the appropriate link

`.HTML` files are preferred but not allowed by all platforms.

<br>

## Before You Submit

Remember to ensure the following before submitting your assignment.

1. Name your files using this format: **lab-##-lastname.rmd** and **lab-##-lastname.html**
2. Show both the solution for your code and write out your answers in the body text 
3. Do not show excessive output; truncate your output, e.g. with function `head()`
4. Follow appropriate styling conventions, e.g. spaces after commas, etc.
5. Above all, ensure that your conventions are consistent

See [Google's R Style Guide](https://google.github.io/styleguide/Rguide.xml) for examples of common conventions.

<br>
<br>

## Common Knitting Issues

`.RMD` files are knit into `.HTML` and other formats procedural, or line-by-line.

* An error in code when knitting will halt the process; error messages will tell you the specific line with the error
* Certain functions like `install.packages()` or `setwd()` are bound to cause errors in knitting 
* Altering a dataset or variable in one chunk will affect their use in all later chunks
* If an object is "not found", make sure it was created or loaded with `library()` in a previous chunk

**If All Else Fails:** If you cannot determine and fix the errors in a code chunk that's preventing you from knitting your document, add `eval = FALSE` inside the brackets of `{r}` at the beginning of a chunk to ensure that R does not attempt to evaluate it, that is: `{r eval = FALSE}`. This will prevent an erroneous chunk of code from halting the knitting process.  

<br>
<br>

<style>

blockquote {
    padding: 11px 22px;
    margin: 0 0 22px;
    font-size: 18px;
    border-left: 5px solid lightgray;
}
</style>



