---
title: "Binding"
author: "Data Science for Public Service"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
---

```{r include = FALSE}

setwd("~/DS4PS/New Textbook Sections")

knitr::opts_chunk$set(echo = FALSE)

library(scales)
library(plotly)
library(pander)
library(tidyverse)

```

<br>

# Defining Binds

Binding is a manner of combining two `data.frame` objects by simply pushing them 
together. 

Suppose we have the following two data frames:

```{r}

abc <- data.frame(Letters = c("A", "B", "C"),
                  Numbers = c(1, 2, 3))

def <- data.frame(Letters = c("D", "E", "F"),
                  Numbers = c(4, 5, 6))
```

<br>

```{r}

abc %>% pander()

```

<br>

```{r}

def %>% pander()

```

<br>
<br>

## Row Binds

If we were to **bind** these data frames by their **rows**, we would stack the 
three rows on top of the other three rows like bricks!

<br>

```{r}

bind_rows(abc, def) %>% 
  pander()

```

<br>
<br>

## Column Binds

Alternatively, we may **bind** a  variable by the **columns** of a data frame,
placing them side-by-side with existing columns. 

<br>

```{r}

nato <- tibble(NATO = c("Alpha", 
                        "Bravo", 
                        "Charlie", 
                        "Delta", 
                        "Echo", 
                        "Foxtrot"))

bind_rows(abc, def) %>% 
  bind_cols(nato) %>% 
  pander()

```

<br>
<br>

# Binding Functions

Binding functions are few and far between. 

Base R functions `rbind()` and `cbind()` are for row binds and column binds, 
respectively.  

Package `dplyr` functions `bind_rows()` and `bind_cols()` are their Tidyverse 
counterparts. 

<br>
<br>

## Use Requirements

You need only pass two or more `data.frame` objects as arguments to any of the 
above functions, with `function(data_frame_1, data_frame_2)` syntax. 

Remember to choose the right function by looking at your data and whether you 
are binding by rows or columns. 

Most importantly, the two `data.frame` objects you are binding must have:

1. Matching or coercible variable classes when joining by rows
2. Matching number of rows if binding by row, and columns if by column

<br>
<br>

## Examples

Recall the above `data.frame` objects represented, respectively, by objects 
`abc` and `def`:

<br>

```{r}

abc %>% pander()

```

<br>

```{r}

def %>% pander()

```

<br>

We can **bind** these data frames by **row** using `rbind()`.

<br>

```{r echo = TRUE, eval = FALSE}

rbind(abc, 
      def)

```

```{r echo = FALSE}

rbind(abc, 
      def) %>% 
  pander()

```

<br>

Similarly, we can use `dplyr` function `bind_rows()`.

<br>

```{r echo = TRUE, eval = FALSE}

library(dplyr)

bind_rows(abc, 
          def)

```

```{r echo = FALSE}

bind_rows(abc, 
          def) %>% 
  pander()

```

<br>
<br>

# Further Resources

Visit documentation for Base R binding functions here:

* Function `r`bind()` Documentation (**[RDocumentation](https://www.rdocumentation.org/packages/R6Frame/versions/0.1.0/topics/rbind)**)
* Function `cbind()` Documentation (**[RDocumentation](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cbind)**)

<br>

Visit documentation for package `dplyr` binding functions here:

* Function `bind_rows()` & `bind_cols()` Documentation (**[Tidyverse.org](https://dplyr.tidyverse.org/reference/bind.html)**)

<br>
<br>
